---
title: "Gov 50 Final Project"
author: "Aomi Miura"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to the political polarization in the US.
Ideally, I would like to investigate how people endorsing right/left-leaning media (newspapers, TV, etc.) have a tendency to reinforce their ideological notions. However, individuals' previously-held ideological notions would be a confounder of the possible causality. Therefore, I wish to adopt "Difference-in-difference" research designs.

##Milestone 2

Does biased media influence people's voting behavior, and if so, does the extent of this influence vary based on race?
In this study, I plan to examine the impact of the introduction of Fox News into the television market in the 1990s on the increase in Republican votes share in Presidential elections in selected towns.

I hypothesize that (1) the introduction of FOX News increased the Republican vote share in presidential elections, and (2) the extent of the increase in Republican vote share was more modest in towns with a higher proportion of black population compared to those with a comparatively lower proportion of black population (Hypothesis (2) is based on the premise that Hypothesis (1) will be supported).

Recent behavioral and cognitive studies suggest that voters can be subjected to the media. However, given that African American people are generally not strong supporters of the Republican party and are often against Republican policies in general, the effects of media persuasion may be limited in towns with a predominantly black population.

The dataset I am going to use is  "Fox News Rollout." (https://www.dropbox.com/s/i1n16wng61tq02x/foxnews.csv?dl=0)

There will be two independent variables: the first variable will indicate whether Fox News was introduced as of 2000, coded as 1 for Fox News already being introduced and 0 for Fox News not being introduced yet. The second variable will represent the proportion of black population in each targeted town in 1990. I will set a suitable threshold to distinguish towns with a relatively higher black population from those with a lower proportion.
My dependent variable is the two-party votes share for Republicans in the 1996 and 2000 Presidential elections. For example, if the difference between Republican vote shares in 1996 and 2000 is positive, it indicates an increase in Republican vote share in the town.

If I observe an increase in Republican vote share from the 1996 to the 2000 election in towns where Fox News was available, this would support my Hypothesis (1). On the other hand, if I observe little to no change or a decrease in Republican vote share from 1996 to 2000 in towns with access to Fox News, my Hypothesis (1) would be disproved. Furthermore, if the increase in Republican vote share in black-dominated towns is not as high as that observed in Hypothesis (1), my Hypothesis (2) would be supported. Conversely, if the extent of the increase is similar to that observed in Hypothesis (1), my Hypothesis (2) would be disproved

##Milestone 3

```{r}
library(tidyverse)
data <- read_csv("foxnews.csv")
data
```

```{r}
state_prop_table <- data |>
  group_by(state) |>
  summarize(count= n())|>
  mutate(prop = count / sum(count))|>
  select(-count)

state_prop_table
```

```{r}
state_prop <- state_prop_table|>
  slice_max(prop, n = 10) |>
  ggplot(aes(x = prop, y = fct_reorder(state, prop))) +
  geom_col()+
  labs(x= "Proportion of Each State included in Data",
       y = "The Name of States")

state_prop
```

##Milestone 4
```{r}
data <- data |>
  mutate(treatment = if_else(subrf2000 == 1, "Treatment", "Control")) |>
  filter(treatment == "Treatment" | treatment == "Control") |>
  mutate(change_1996_2000 = gopvoteshare2000 - gopvoteshare1996)

avg_change <- data |>
  summarize(mean_1996_2000 = mean(change_1996_2000, na.rm = TRUE))

avg_change
```

```{r}
DID <- data |>
  group_by(treatment) |>
  summarize(avg_change = mean(vote_change)) |>
  pivot_wider(names_from = treatment,
              values_from = avg_change) |>
  mutate(DID = Treatment - Control)

DID
```
The first hypothesis of the research was that the introduction of FOX News increased the Republican vote share in presidential elections.
However, difference-in-difference research conducted above indicate that the hypothesis has been disapproved.

In this context, treatment group are towns where FOX news were already introduced as of 2000, while control group are those where FOX did not enter yet. 
The average change of the DOP vote shares between 1996 and 2000 election in FOX-available towns were 0.04277223, whereas that in FOX-unavailable towns were 0.06888928, indicating that the introduction of FOX news in the town actually decreased the DOP vote share. This could potentially suggest that the entry of FOX news in the town have backfired and alliniated the voters away from Republican support.








